---
title: "DAM Prices Retrieval and Analysis"
format: 
  html:
    toc: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{DAM Prices Retrieval and Analysis}
  %\VignetteEngine{quarto::html}
---

## Introduction

This vignette demonstrates how to retrieve and analyze Day-Ahead Market (DAM) prices for different countries and date ranges using the `api_entsoe_dam_prices()` and `entsoe_dam_prices()` functions. We will cover:
1. Retrieving raw DAM prices data using an API method.
2. Analyzing and plotting DAM prices for a given country and date range.

## Setup

The first step is to load the necessary packages. We use `data.table` for efficient data manipulation, `magrittr` for piping, `ggplot2` for plotting, and a custom package `fluxer` for API integration.

```{r setup}
box::use(data.table[...])
box::use(fluxer[...])
```

## 1. API Method for Retrieving DAM Prices
### 1.1 Raw DAM Prices
We start by using the api_entsoe_dam_prices function to retrieve raw DAM prices for a specific country and date range. For this example, we retrieve data for the Czech Republic from January 4, 2024, to January 6, 2024.


```{r}
country <- "Czech Republic"
from_data <- '2024-01-04'
to_data <- '2024-01-06'
# api_key = Sys.getenv('ENTSOE_KEY')  # Uncomment and provide API key if needed

dam_prices <- api_entsoe_dam_prices(country, from_data, to_data)
print(dam_prices)

```

The function api_entsoe_dam_prices takes the country, start date, and end date as arguments and returns raw DAM prices data. The raw data is printed for inspection.

## 1.2 Processed DAM Prices
Next, we retrieve and process the DAM prices for Spain over a longer period, from January 4, 2024, to February 8, 2024, using the entsoe_dam_prices function. This function allows us to process the raw data and optionally visualize it.


```{r}
country <- "Spain"
from_data <- '2024-01-04'
to_data <- '2024-02-08'
# api_key = Sys.getenv('ENTSOE_KEY')  # Uncomment and provide API key if needed

dam_prices <- entsoe_dam_prices(country, from_data, to_data, verbose = TRUE, plot = TRUE)
print(dam_prices)

```

In this step, the `entsoe_dam_prices()` function not only retrieves the data but also processes it and creates a plot if the `plot` argument is set to `TRUE`.

## Conclusion
In this vignette, we demonstrated how to retrieve raw and processed DAM prices for different countries using the `api_entsoe_dam_prices()` and `entsoe_dam_prices()` functions.

Analyze the data by plotting DAM prices over a specific date range.

You can modify the parameters such as country, start date, and end date to retrieve and analyze DAM prices for other regions and periods.

Make sure to handle API keys securely by using environment variables (`Sys.getenv('ENTSOE_KEY')`), especially when deploying in production environments.

# Additional Notes
- The `entsoe_dam_prices()` function includes optional arguments like `verbose` and `plot` for more detailed output and visualizations.  
- If you are working with large datasets, consider using `data.table` functions for efficient data manipulation.
