---
title: "GME Prices Data Processing"
format: 
  html:
    toc: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{GME Prices Data Processing}
  %\VignetteEngine{quarto::html}
---

## Introduction

This vignette demonstrates how to download and process GME (Gestore dei Mercati Energetici) price data for both MGP (Mercato Giornaliero di Produzione) and MSD (Mercato dei Servizi di Dispacciamento). The data is retrieved from a remote FTP server and processed into clean datasets.

The code is divided into two parts:
1. **MGP Prices**: Download and process MGP prices.
2. **MSD Prices**: Download and process MSD prices.

## Setup

We first define the parameters required for downloading and processing the files. The data types, FTP credentials, and output directory are set as follows:

```{r setup, eval=FALSE}
# Set the parameters for both MGP and MSD
n <- 2  # Number of files to process

data_type_mgp <- 'MGP'
data_type_msd <- 'MSD'

username <- "PIASARACENO"
password <- "18N15C9R"
output_dir <- "inst/data"
```

## 1. MGP Prices
### 1.1 Get Available Files
We use the function `gme_prices_get_files()` to list all available MGP price files in the GME folder. We then select the last n files for processing.


```{r, eval=FALSE}
library(data.table)
library(fluxer)
# Get MGP prices files available at GME folder
gme_mgp_prices_files <- gme_prices_get_files(data_type = data_type_mgp, 
                                             output_dir = output_dir, 
                                             username = username, 
                                             password = password)

# Select the last n files
last_n_files_mgp <- tail(gme_mgp_prices_files, n)
print(last_n_files_mgp)

```

## 1.2 Download and Clean Data
We then download and process the data from the selected MGP price files using the function `gme_download_prices_file()`. The downloaded data is cleaned and combined into a single dataset.

```{r, eval=FALSE}
list_mgp_prices <- lapply(last_n_files_mgp, function(file) {
    tryCatch({
        # Call gme_download_prices_file with explicit arguments
        gme_download_prices_file(
            filename = file,
            data_type = data_type_mgp,
            output_dir = output_dir,
            username = username,
            password = password,
            raw = FALSE
        )
    }, error = function(e) {
        # In case of an error, return NULL
        message("Error processing file: ", file, " - ", e$message)
        return(NULL)
    })
})

# Combine the downloaded MGP prices into a single data.table
dt_mgp_prices <- rbindlist(list_mgp_prices, fill = TRUE)
print(dt_mgp_prices)

```

## Conclusion
This vignette demonstrated how to download and process GME prices data for both MGP and MSD. By using FTP credentials and defining the necessary parameters, the data is cleaned and combined into two datasets: one for MGP prices and one for MSD prices.

You can customize the number of files to process by adjusting the parameter n and modify the data processing functions to meet your specific needs.

Don't forget to handle errors gracefully in case any file fails to download or process.

## Additional Notes
## Additional Notes
- Make sure to replace the FTP `username` and `password` with your actual credentials.  
- The data is saved to the directory defined by `output_dir`, which is `inst/data` by default.  
- You can modify the `gme_download_prices_file()` function to include additional functionality or error handling if needed.  
